---
title: "Part 1: Time series wrangling & forecasting"
author: "Larissa Neilson"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE)

library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
library(broom)
```

### Read in energy data and convert to a tsibble

```{r}
energy <- read_csv(here("data", "energy.csv"))

# Wrangle data to get proper date column
energy_ts <- energy %>% 
  mutate(date = yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)
```

### Exploratory time series visualization

#### Raw data ggplot

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = "Residential energy consumption \n (Trillion BTU)")
```

- Overall increasing trend, but stability (and possibly a slight decreasing trend) starting around 2005
- Clear seasonality, with a dominant seasonal feature and also a secondary peak each year - that secondary peak has increased substantially
- No notable cyclicality or outliers

#### Seasonplot

```{r}
energy_ts %>% 
  gg_season(y = res_total) +
  theme_minimal() +
  labs(x = "month", y = "residential energy consumption (trillion BTU)")
```

- The highest residential energy usage is around December / January / February
- There is a secondary peak around July & August (that’s the repeated secondary peak we see in the original time series graph)
- We can also see that the prevalence of that second peak has been increasing over the course of the time series: in 1973 (orange) there was hardly any summer peak. In more recent years (blue/magenta) that peak is much more prominent.

#### Subseries plot

```{r}
energy_ts %>% 
  gg_subseries(res_total)
```

- There is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.

#### Decomposition (by STL)

```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~  season()))

# Visualize the decomposed components
components(dcmp) %>% 
  autoplot() +
  theme_minimal()
```

#### Autocorrelation function (ACF)

```{r}
energy_ts %>% 
  ACF(res_total) %>%
  autoplot()
```

- Observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.

### Forecasting by Holt-Winters exponential smoothing

```{r}
# Create the model:
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# Forecast using the model 10 years into the future:
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

# Plot just the forecasted values (with 80 & 95% CIs):
energy_forecast %>% 
  autoplot()

# Or plot it added to the original data:
energy_forecast %>% 
  autoplot(energy_ts)
```   

#### Assessing residuals

```{r}
# Append the predicted values (and residuals) to original energy data
energy_predicted <- augment(energy_fit)

# Now plot the actual energy values (res_total), and the predicted values (stored as .fitted) atop them:
ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total)) +
  geom_line(aes(x = date, y = .fitted, color = "pink"))
```

Explore the residuals  next.
Some important considerations: Residuals should be uncorrelated, centered at 0, and ideally normally distributed. One way we can check the distribution is with a histogram:

```{r}
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()
```

Relatively normally distributed, and centered at 0 (we could find summary statistics beyond this to further explore)

### Other forecasting methods

ETS forecasting, seasonal naive (SNAIVE) and autoregressive integrated moving average (ARIMA)

```{r}
# Fit 3 different forecasting models (ETS, ARIMA, SNAIVE):
energy_fit_multi <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M")),
    arima = ARIMA(res_total),
    snaive = SNAIVE(res_total)
  )

# Forecast 3 years into the future (from data end date)
multi_forecast <- energy_fit_multi %>% 
  forecast(h = "3 years")

# Plot the 3 forecasts
multi_forecast %>% 
  autoplot(energy_ts)

# Or just view the  forecasts (not the similarity across models):
multi_forecast %>% 
  autoplot()
```

